---
title: "EDA of Streamflow Data"
author: "Mengran Li"
date: "2021/2/22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data preparation

library packages we need.
```{r}
# Big data processing
library(data.table)
# ggplot, dplyr, etc.
library(tidyverse)
# Smooth
library(KernSmooth)
# Map
library(maps)
# Animated Graphics
library(gganimate)
```

Load data.
```{r}
load("Siteinfo.Rdata")
load("Streamflow.Rdata")

gauge <- read.table("gauge_information.txt", header = FALSE)
names(gauge) <- gauge[1, ]
gauge <- gauge[-1, ]

Region_name <- read.table("Region_name.txt", header = FALSE, sep=",")
names(Region_name) <- Region_name[1, ]
Region_name <- Region_name[-1, ]
```

-  Delete data of 2020-01-01
-  Add Site information into the dataset
-  Round down the year, as DecYear is the original date transformed into decimal year.
   For instance, "1980-01-01" is 1980.000.
-  Transform the list as dataframe (a special data table).
```{r}
for (i in 1:length(Streamflow)) {
  Streamflow[[i]]$Region <- gauge[i, 1]
  Streamflow[[i]]$Site <- gauge[i, 2]
}

Stream <- rbindlist(Streamflow)[
  ,DecYear := floor(DecYear)][
    DecYear != 2020]
```

But I notice that there are some observations of Q below 0.
Their qualifier is "A:e" (77W+), which means they are estimated by the USGS as the actual streamflow. It may be not proper to remove all the estimated values. I delete the negtive estimations (about 7000+), and remain the positive estimations as actual streamflow. If the treatment is not suitable or you have any other method, please let me know.






Select variables we need.
```{r}
Stream_plot <- Stream[, c("Region", "Site", "DecYear", "Month", "Q")]
head(Stream_plot)
```

List of Variables:

|Region|Site|DecYear|Month|Q|
|:---:|:---:|:---:|:---:|:---:|
|Hydrology Region|Observation Site|Calendar Year|Month|Daily Streamflow|
|18 Regions|671 Sites|1980-2019,40 years|Jan-Dec,12 months|$m^3/s$|

There are totally 9,559,629 observations.

## Temporal EDA

### Distribution of all data for every year.

#### All Regions
```{r, eval = FALSE}
blank = rep(" ",4)
label = c()
for (i in 1:8){
  label = c(label, 1980 + 5* (i-1), blank)
}
Stream_plot[,DecYear:=as.factor(DecYear)] %>%
  ggplot(aes(x = DecYear, y = Q)) +
    geom_boxplot()+
    scale_x_discrete(labels = label, breaks=1980:2019)
```

#### Each region

```{r, eval = FALSE}
blank = rep(" ",9)
label = c()
for (i in 1:4){
  label = c(label, 1980 + 10* (i-1), blank)
}
Stream_plot[,DecYear:=as.factor(DecYear)] %>%
  ggplot(aes(x = DecYear, y = Q, colour = Region)) +
  geom_boxplot() +
  scale_x_discrete(labels = label, breaks=1980:2019)+
  facet_wrap(~Region)
```

Rendering takes a so long time so that rmd doesn't work. I have saved the figure as Figure2.

### Trend detection of monthly data over years

```{r}
Stream_Series <- Stream_plot[, mean(Q, na.rm = TRUE), 
                             by = .(DecYear, Month)] 
Stream_Series%>%
       spread(key="Month", value="V1")
```

Smooth the lines with $locpoly()$

```{r}
H <- as.matrix(Stream_Series[,dpill(DecYear, V1), by = Month])

Stream_Series_fit <- list()
for (i in 1:12){
  data <- Stream_Series[Month==i]
  Year <- data$DecYear
  Q <- data$V1
  Stream_Series_fit[[i]] <- locpoly(Year, Q, bandwidth = H[i,2])
  Stream_Series_fit[[i]]$Month <- i
}

rbindlist(Stream_Series_fit)[
  ,Month:=factor(Month, labels = month.abb)] %>%
  ggplot(aes(x=x,y=y,group=Month,colour=Month))+
  geom_line()+
  labs(x = "Year", y = "Streamflow", title = "Trend of Monthly Streamflow")+
  coord_fixed(ratio=1)
```


However it looks messy. Given by the dry season and rainy season, it is possible to exist different trends among months. Thus I try to divide the months into two groups.

```{r}
Season <- rbindlist(Stream_Series_fit)
Season[which(Season$Month%in% c(3:5, 8:10)),"Group"] <- "A"
Season[which(!Season$Month%in% c(3:5, 8:10)),"Group"] <- "B"
Season$Month <- factor(Season$Month, labels = month.abb)

Season %>%
  ggplot(aes(x=x,y=y))+
  geom_line(aes(group=Month,colour=Month))+
  labs(x = "Year", y = "Streamflow", title = "Grouped Trend of Monthly Streamflow")+
  facet_wrap(~Group)+
  coord_fixed(ratio=1)

```
Group A contains Mar, Apr, May, Aug, Sep and Oct. They all looks stable.
Group B includes the others, which fluctuate periodically.


## Spatial EDA

### Statistics for all locations

```{r}
Stream_summary <-  Stream[, .(
  mean(Q, na.rm = TRUE), 
  sd(Q, na.rm = TRUE),
  quantile(Q, 0.05, na.rm = TRUE),
  quantile(Q, 0.1, na.rm = TRUE),
  quantile(Q, 0.2, na.rm = TRUE),
  quantile(Q, 0.5, na.rm = TRUE),
  quantile(Q, 0.7, na.rm = TRUE),
  quantile(Q, 0.9, na.rm = TRUE),
  quantile(Q, 0.95, na.rm = TRUE),
  quantile(Q, 0.975, na.rm = TRUE),
  quantile(Q, 0.99, na.rm = TRUE))]
names(Stream_summary) <- c("Mean", "Sd",
                       paste0("q", c(0.05, 0.1, 0.2, 0.5, 0.7, 0.9, 0.95, 0.975, 0.99)))
Stream_summary
```

```{r}
map_site <- Siteinfo[, c("site_no", "dec_lat_va", "dec_lon_va")]
names(map_site) <- c("Site", "lat", "long")

Stream_map_full <- full_join(Stream_plot, map_site, by = c("Site" = "Site"))
```

### Mean Exceedance

$$u=\hat{F}^{-1}(q)$$
Hence,
-  Assign $u=\hat{F}^{-1}(q)$
-  Select $Q > u$ (right tail), $Q < u$ (left tail)
-  Calculate mean of $Q$

### Right tail

#### q0.7

```{r}
u <- Stream_summary[,q0.7]
Stream_map_full[Q > u][
  , MeanExceedance:=mean(Q, na.rm = TRUE), by=Site]%>%
  ggplot(aes(long, lat)) +
  geom_point(aes(color = MeanExceedance, size = I(3), alpha = I(0.5))) +
  scale_colour_gradientn(colours = c("purple", "skyblue","green","yellow","orange","red")) +
  labs(title = "Mean Exceedance of 70% quantile")+
  borders("state")+
  coord_fixed(ratio=1.5)
```

#### q0.9

```{r}
u <- Stream_summary[,q0.9]
Stream_map_full[Q > u][
  , MeanExceedance:=mean(Q, na.rm = TRUE), by=Site]%>%
  ggplot(aes(long, lat)) +
  geom_point(aes(color = MeanExceedance, size = I(3), alpha = I(0.5))) +
  scale_colour_gradientn(colours = c("purple", "skyblue","green","yellow","orange","red")) +
  labs(title = "Mean Exceedance of 90% quantile")+
  borders("state")+
  coord_fixed(ratio=1.5)
```

#### q0.95

```{r}
u <- Stream_summary[,q0.95]
Stream_map_full[Q > u][
  , MeanExceedance:=mean(Q, na.rm = TRUE), by=Site]%>%
  ggplot(aes(long, lat)) +
  geom_point(aes(color = MeanExceedance, size = I(3), alpha = I(0.5))) +
  scale_colour_gradientn(colours = c("purple", "skyblue","green","yellow","orange","red")) +
  labs(title = "Mean Exceedance of 95% quantile")+
  borders("state")+
  coord_fixed(ratio=1.5)
```

#### q0.975

```{r}

u <- Stream_summary[,q0.975]
Stream_map_full[Q > u][
  , MeanExceedance:=mean(Q, na.rm = TRUE), by=Site]%>%
  ggplot(aes(long, lat)) +
  geom_point(aes(color = MeanExceedance, size = I(3), alpha = I(0.5))) +
  scale_colour_gradientn(colours = c("purple", "skyblue","green","yellow","orange","red")) +
  labs(title = "Mean Exceedance of 97.5% quantile")+
  borders("state")+
  coord_fixed(ratio=1.5)
```

#### 0.99

```{r}
u <- Stream_summary[,q0.99]
Stream_map_full[Q > u][
  , MeanExceedance:=mean(Q, na.rm = TRUE), by=.(Site, DecYear)]%>%
  ggplot(aes(long, lat)) +
  geom_point(aes(color = MeanExceedance, size = I(3), alpha = I(0.5))) +
  scale_colour_gradientn(colours = c("purple", "skyblue","green","yellow","orange","red")) +
  labs(title = "Mean Exceedance of 99% quantile")+
  borders("state")+
  coord_fixed(ratio=1.5)
```




### Left Tail

#### q0.2

```{r}
u <- Stream_summary[,q0.2]
Stream_map_full[Q < u][
  , MeanExceedance:=mean(Q, na.rm = TRUE), by=Site]%>%
  ggplot(aes(long, lat)) +
  geom_point(aes(color = MeanExceedance, size = I(3), alpha = I(0.5))) +
  scale_colour_gradientn(colours = c("purple", "skyblue","green","yellow","orange","red")) +
  labs(title = "Mean Exceedance of 20% quantile")+
  borders("state")+
  coord_fixed(ratio=1.5)
```

#### q0.1

```{r}
u <- Stream_summary[,q0.1]
Stream_map_full[Q < u][
  , MeanExceedance:=mean(Q, na.rm = TRUE), by=Site]%>%
  ggplot(aes(long, lat)) +
  geom_point(aes(color = MeanExceedance, size = I(3), alpha = I(0.5))) +
  scale_colour_gradientn(colours = c("purple", "skyblue","green","yellow","orange","red")) +
  labs(title = "Mean Exceedance of 10% quantile")+
  borders("state")+
  coord_fixed(ratio=1.5)
```

#### q0.05

```{r}
u <- Stream_summary[,q0.05]
Stream_map_full[Q < u][
  , MeanExceedance:=mean(Q, na.rm = TRUE), by=Site]%>%
  ggplot(aes(long, lat)) +
  geom_point(aes(color = MeanExceedance, size = I(3), alpha = I(0.5))) +
  scale_colour_gradientn(colours = c("purple", "skyblue","green","yellow","orange","red")) +
  labs(title = "Mean Exceedance of 5% quantile")+
  borders("state")+
  coord_fixed(ratio=1.5)
```

