---
title: "Untitled"
author: "Mengran Li"
date: "2021/3/1"
output: html_document

---

## Questions in the note

###  <font size="5"> Do we have data for almost every day within the months?</font>

<font size="4">There is $40\times365=14600$ days. Therefore we should have $671\times14600=9796600$ observations. Acturally we have $9559629$ observations, in the condition of deleting some data. The propotion is $9559629/9796600\times100\%=97.58\%$, which is acceptable in my view.</font>

###  <font size="5"> Are there fewer rivers/stream in the plateau area?</font>

```{r, echo = FALSE, message = FALSE}
library(tidyverse)
library(leaflet)
setwd("F:\\StreamFlow")
load("Stream.Rdata")

```

```{r, echo = FALSE}
GetURL <- function(service, host = "basemap.nationalmap.gov") {
  sprintf("https://%s/arcgis/services/%s/MapServer/WmsServer", host, service)
}
Siteinfo %>%
  leaflet() %>%
  addTiles() %>%
  addCircleMarkers(
    lng = ~dec_lon_va,
    lat = ~dec_lat_va,
    radius = ~3,
    stroke = FALSE,
    fill = TRUE,
    color = "red",
    fillOpacity = 0.4,
    group = "Site"
  )%>%
    addCircleMarkers(
    lng = ~dec_lon_va,
    lat = ~dec_lat_va,
    radius = ~1,
    stroke = FALSE,
    fill = TRUE,
    fillOpacity = 0,
    group = "No Site"
  )%>%
  addWMSTiles(GetURL("USGSHydroCached"),layers = "0")%>%
  addLayersControl(
          overlayGroups =c("Site", "No Site"),
          options = layersControlOptions(collapsed=FALSE)
          )
  
```
<font size="4">
The base plot is from USGS, [Hydro-NHD - A base map that highlights hydrographic and NHD features](https://waterdata.usgs.gov/blog/basemaps/).
It is obvious that there are fewer both rivers and observation sites in the plateau area. Click on the button, you can see the river map with/ without observation sites.</font>


